---
title: "IDS investigation worksheet"
author: "by Team-Name: User1, User2, User3, User4 & User5"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(tidymodels)
# Add any other libraries here



```


```{r load-data}
# load your data 
df <- read_csv("data/AdjudicationsQ12023.csv")
view(df)

  

```

```{r tidy-data}
prisondata <- df[-c(1, 2, 5, 11)]

prisondata <- prisondata %>%
  rename("prison_type" = "Predominant function of establishment", "establishment" = "Establishment", "sex" = "Sex", "age_group" = "Age group", "ethnicity" = "Ethnicity", "religion" = "Religion", "offence" = "Offence", "number_of_offences" = "Count")

prisondata <- prisondata %>% 
mutate(violent = case_when(offence == "Violence" ~ "Yes", offence == "Wilful Damage" ~ "Yes", TRUE ~ "No")) %>%
mutate(sex = case_when(sex == "M" ~ 1,
                       sex == "F" ~ 2))

prisondata <- prisondata %>% mutate(reoffender = case_when(number_of_offences == 1 ~ 1, number_of_offences != 1 ~ 0))

view(prisondata)

prisondata %>%
  group_by(age_group) %>%
  count()

```
```{r}
ggplot(data = prisondata, mapping = aes(x = age_group, fill = violent)) +
  geom_bar() +
  facet_wrap(~prison_type, ncol = 2, scales = "free") +
  theme(
    axis.text = element_text(size = 5)
)
```

```{r}
#summary statistics
prisondata%>%
 filter(age_group=="18 - 20")%>%summarise(mean_for_age=mean(number_of_offences,na.rm=TRUE))

prisondata%>%
  filter(age_group=="21 - 24")%>%summarise(mean_for_age=mean(number_of_offences,na.rm=TRUE))

prisondata%>%
  filter(age_group=="25 - 29")%>%summarise(mean_for_age=mean(number_of_offences,na.rm=TRUE))

prisondata%>%
  filter(age_group=="30 - 39")%>%summarise(mean_for_age=mean(number_of_offences,na.rm=TRUE))

prisondata%>%
  filter(age_group=="40 - 49")%>%summarise(mean_for_age=mean(number_of_offences,na.rm=TRUE))

prisondata%>%
  filter(age_group=="50 - 59")%>%summarise(mean_for_age=mean(number_of_offences,na.rm=TRUE))

prisondata%>%
  filter(age_group=="70 and over")%>%summarise(mean_for_age=mean(number_of_offences,na.rm=TRUE))
```

```{r factor-relevelling}
#Relevel prison type

prisondatamodel <- prisondata
prisondatamodel <- prisondatamodel %>% 
  mutate(prison_type = as.factor(prison_type)) %>%
  mutate(prison_type = fct_relevel(prison_type, c("Category A (High Security)", "Category B", "Category C", "Category D (Open)", "Local", "Young Offender Institution (YOI)", "Youth Justice Board", "Female")))

#Relevel sex

prisondatamodel <- prisondatamodel %>%
  mutate(sex = as.factor(sex)) %>%
  mutate(sex = fct_relevel(sex, c("1", "2")))

#Relevel age group

prisondatamodel <- prisondatamodel %>%
  mutate(age_group = as.factor(age_group)) %>%
  mutate(age_group = fct_relevel(age_group, c("15 - 17", "18 - 20", "21 - 24", "25 - 29", "30 - 39", "40 - 49", "50 - 59", "60 - 69", "70 and over")))

#Relevel religion

prisondatamodel <- prisondatamodel %>% 
  filter(religion != "Not recorded") %>%
  mutate(religion = as.factor(religion)) %>%
  mutate(religion = fct_relevel(religion, c("a Christian", "b Muslim", "c Hindu", "d Sikh", "e Buddhist", "f Jewish", "g Other religious groups", "i No religion")))

#Relevel violent

prisondatamodel <- prisondatamodel %>% 
  mutate(violent = as.factor(violent)) %>%
  mutate(violent = fct_relevel(violent, c("Yes", "No")))

#Relevel reoffender

prisondatamodel <- prisondatamodel %>% 
  mutate(reoffender = as.factor(reoffender)) %>%
  mutate(reoffender = fct_relevel(reoffender, c("0", "1")))

prisonmodelsdata <- prisondatamodel
```

```{r splitting data}

pmd_split <- initial_split(prisonmodelsdata, prop = 0.75)

pmd_train <- training(pmd_split)
pmd_test <- testing(pmd_split)
```

```{r building model 1}

pmd_model1 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe1 <- recipe(reoffender ~ violent, data = pmd_train) %>%
  step_dummy(violent, -reoffender)

pmd_workflow1 <- workflow() %>%
  add_model(pmd_model1) %>%
  add_recipe(pmd_recipe1)

pmd_fit1 <- pmd_workflow1 %>%
  fit(data = pmd_train)
```

```{r building model 2}

pmd_model2 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe2 <- recipe(reoffender ~ violent + age_group + sex + prison_type + religion, data = pmd_train) %>%
  step_dummy(all_nominal_predictors(), -reoffender)

pmd_workflow2 <- workflow() %>%
  add_model(pmd_model2) %>%
  add_recipe(pmd_recipe2)

pmd_fit2 <- pmd_workflow2 %>%
  fit(data = pmd_train)
```
```{r}
tidy(pmd_fit1)
```


```{r}
tidy(pmd_fit2)
```


```{r pred-roc for models}

pmd_prediction1 <- predict(pmd_fit1, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_prediction2 <- predict(pmd_fit2, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_roccurve1 <- pmd_prediction1 %>%
  roc_curve(truth = reoffender, .pred_1) %>% autoplot()

pmd_rocauc1 <- pmd_prediction1 %>%
  roc_auc(truth = reoffender, .pred_1)

pmd_roccurve2 <- pmd_prediction2 %>%
  roc_curve(truth = reoffender, .pred_1) %>% autoplot()

pmd_rocauc2 <- pmd_prediction2 %>%
  roc_auc(truth = reoffender, .pred_1)
```

```{r}
#looking at predictions
pmd_prediction1 <- predict(pmd_fit1, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_prediction1
```

```{r}
#roc curve for prediction 1
pmd_prediction1 %>%
  roc_curve(truth = reoffender, .pred_1) %>%
  autoplot()
```

```{r}
#looking at predictions 2
pmd_prediction2 <- predict(pmd_fit2, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_prediction2
```
```{r}
#roc curve for prediction 2
pmd_prediction2 %>%
  roc_curve(truth = reoffender, .pred_1) %>%
  autoplot()

```

```{r}

pmd_model3 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe3 <- recipe(reoffender ~ prison_type, data = pmd_train) %>%
  step_dummy(all_nominal_predictors(), -reoffender)

pmd_workflow3 <- workflow() %>%
  add_model(pmd_model3) %>%
  add_recipe(pmd_recipe3)

pmd_fit3 <- pmd_workflow3 %>%
  fit(data = pmd_train)
```

```{r}
tidy(pmd_fit3)
```

```{r}
pmd_prediction3 <- predict(pmd_fit3, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_prediction3
```


```{r}
#roc curve for prediction 3, prison type
pmd_prediction3 %>%
  roc_curve(truth = reoffender, .pred_1) %>%
  autoplot()
```

```{r}
#area under the curve for prediction 3
pmd_prediction3 %>%
  roc_auc (truth = reoffender, .pred_1)
```

```{r}
pmd_model4 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe4 <- recipe(violent ~ sex, data = pmd_train) %>%
  step_dummy(all_nominal_predictors(), -violent)

pmd_workflow4 <- workflow() %>%
  add_model(pmd_model4) %>%
  add_recipe(pmd_recipe4)

pmd_fit4 <- pmd_workflow4 %>%
  fit(data = pmd_train)
```

```{r}
pmd_prediction4 <- predict(pmd_fit4, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_prediction4
```


```{r}
pmd_prediction4 %>%
  roc_curve(truth = violent, .pred_Yes) %>%
  autoplot()
```
```{r}
pmd_prediction4 %>%
  roc_auc (truth = violent, .pred_Yes)
```

