---
title: "Crimes Committed within Prisons in the UK"
subtitle: "correlation between violence and re-offending"
author: "The Ravens <br> Raven, Elanna, Maisie, Julia"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---

```{r install-package, include = FALSE, eval = FALSE}
# Copy and paste the following code into your console to download and install
# the `xaringan` package that contains all of the code which allows you 
# to create presentation slides in Rmarkdown
install.packages('xaringan')
```


```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# Load your data here
df <- read_csv("data/AdjudicationsQ12023.csv")
prisondata <- df[-c(1, 2, 5, 11)]

prisondata <- prisondata %>%
  rename("prison_type" = "Predominant function of establishment", "establishment" = "Establishment", "sex" = "Sex", "age_group" = "Age group", "ethnicity" = "Ethnicity", "religion" = "Religion", "offence" = "Offence", "number_of_offences" = "Count")

prisondata <- prisondata %>% 
mutate(violent = case_when(offence == "Violence" ~ "Yes", offence == "Wilful Damage" ~ "Yes", TRUE ~ "No")) %>%
mutate(sex = case_when(sex == "M" ~ 1,
                       sex == "F" ~ 2))

prisondata <- prisondata %>% mutate(reoffender = case_when(number_of_offences == 1 ~ 0, number_of_offences != 1 ~ 1))

```

```{r include=FALSE}

#Background image
style_xaringan(
  title_slide_background_image = "img/prisonbars.jpg"
)
```

class: inverse, center, middle

## Investigating the correlation between the violence of the prisoner and the number of reoffences committed. 
---
class: centre, middle

## Our Data Set and Research Question
---
# The Prison Data Set

- Our original data set came from the Government website, and details the criminal offences committed in prisons in England and Wales between the months of January and March of 2023.

- We tidied the data, mutating the detailed offence to show whether the crime committed was violent or non-violent. When doing this, we took violent to mean "the use of physical force as to injure, abuse, damage or destroy". 

- The variables in the data set were mainly characters so we changed many of these variables to be numerical to make producing models and visualisations easier. 
---
# Our research question and its importance?

- We wanted to investigate whether any variables of the prisoner like sex, the prison type they are at, violence of the prisoner and age had any correlation with reoffending.

- Our dataset has 24,647 prisoners who committed a crime in prison during the first quarter of 2023. Of these prisoners, 8137 reoffended multiple times. If a trend, and therefore a solution could be found, the number of reoffenders could decrease creating a safer environments in prisons.
---

class: inverse, centre, middle
# Our Findings 

---
# Number of Reoffending Crimes by Prison Type 

.pull-left[
- Prison categories are based on the governments' own categories where category A is highest security, and then B and so on. 
- The violence of the crime in represented by the colours, red being non-violent and blue being violent.
]

.pull-right[
``` {r plot, echo = FALSE}

ggplot(data = prisondata, mapping = aes(x = age_group, fill = violent)) +
  geom_bar() +
  facet_wrap(~prison_type, ncol = 2, scales = "free") +
  theme(
    axis.text = element_text(size = 5))
```
]

---
class: inverse, centre, middle

# Which factors affect the reoffending?

---
#Prison Type and Violence 
.pull-right[
- The area under the ROC curve is 0.474 which tells us that the prison type variable is not effective at predicting whether a prisoner will reoffend.
]
.pull-left[
```{r, echo = FALSE}
prisondatamodel <- prisondata
prisondatamodel <- prisondatamodel %>% 
  mutate(prison_type = as.factor(prison_type)) %>%
  mutate(prison_type = fct_relevel(prison_type, c("Category A (High Security)", "Category B", "Category C", "Category D (Open)", "Local", "Young Offender Institution (YOI)", "Youth Justice Board", "Female")))

prisondatamodel <- prisondatamodel %>%
  mutate(sex = as.factor(sex)) %>%
  mutate(sex = fct_relevel(sex, c("1", "2")))

prisondatamodel <- prisondatamodel %>%
  mutate(age_group = as.factor(age_group)) %>%
  mutate(age_group = fct_relevel(age_group, c("15 - 17", "18 - 20", "21 - 24", "25 - 29", "30 - 39", "40 - 49", "50 - 59", "60 - 69", "70 and over")))

prisondatamodel <- prisondatamodel %>% 
  filter(religion != "Not recorded") %>%
  mutate(religion = as.factor(religion)) %>%
  mutate(religion = fct_relevel(religion, c("a Christian", "b Muslim", "c Hindu", "d Sikh", "e Buddhist", "f Jewish", "g Other religious groups", "i No religion")))

prisondatamodel <- prisondatamodel %>% 
  mutate(violent = as.factor(violent)) %>%
  mutate(violent = fct_relevel(violent, c("Yes", "No")))

prisondatamodel <- prisondatamodel %>% 
  mutate(reoffender = as.factor(reoffender)) %>%
  mutate(reoffender = fct_relevel(reoffender, c("0", "1")))

prisonmodelsdata <- prisondatamodel

pmd_split <- initial_split(prisonmodelsdata, prop = 0.75)

pmd_train <- training(pmd_split)
pmd_test <- testing(pmd_split)

pmd_model3 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe3 <- recipe(reoffender ~ prison_type, data = pmd_train) %>%
  step_dummy(all_nominal_predictors(), -reoffender)

pmd_workflow3 <- workflow() %>%
  add_model(pmd_model3) %>%
  add_recipe(pmd_recipe3)

pmd_fit3 <- pmd_workflow3 %>%
  fit(data = pmd_train)

pmd_prediction3 <- predict(pmd_fit3, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)
```

``` {r building model, echo = FALSE}
pmd_prediction3 %>%
  roc_curve(truth = reoffender, .pred_1) %>%
  autoplot()

```
]
---
# Reoffending and Violence 

.pull-left[
```{r, echo = FALSE}
pmd_model1 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe1 <- recipe(reoffender ~ violent, data = pmd_train) %>%
  step_dummy(violent, -reoffender)

pmd_workflow1 <- workflow() %>%
  add_model(pmd_model1) %>%
  add_recipe(pmd_recipe1)

pmd_fit1 <- pmd_workflow1 %>%
  fit(data = pmd_train)

pmd_prediction1 <- predict(pmd_fit1, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_prediction1 %>%
  roc_curve(truth = reoffender, .pred_1) %>%
  autoplot()

```

]

.pull-right[
- The area under the curve is 0.4479 which shows there is no correlation between the violence of the prisoner and reoffending.
]
---
# Reoffending and violence, age, sex, religion and prison type 

.pull-left[
```{r, echo = FALSE, include = FALSE}
pmd_model2 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe2 <- recipe(reoffender ~ violent + age_group + sex + prison_type + religion, data = pmd_train) %>%
  step_dummy(all_nominal_predictors(), -reoffender)

pmd_workflow2 <- workflow() %>%
  add_model(pmd_model2) %>%
  add_recipe(pmd_recipe2)

pmd_fit2 <- pmd_workflow2 %>%
  fit(data = pmd_train)

pmd_prediction2 <- predict(pmd_fit2, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

```
```{r, echo = FALSE}
pmd_prediction2 %>%
  roc_curve(truth = reoffender, .pred_1) %>%
  autoplot()
```

]

.pull-right[
- The area under the curve is 0.374 which shows there is no correlation between all the variables and reoffending.
]

---
# Violence and Sex

.pull-left[
```{r, echo = FALSE}
pmd_model4 <- logistic_reg() %>%
  set_engine("glm")

pmd_recipe4 <- recipe(violent ~ sex, data = pmd_train) %>%
  step_dummy(all_nominal_predictors(), -violent)

pmd_workflow4 <- workflow() %>%
  add_model(pmd_model4) %>%
  add_recipe(pmd_recipe4)

pmd_fit4 <- pmd_workflow4 %>%
  fit(data = pmd_train)

pmd_prediction4 <- predict(pmd_fit4, pmd_test, type = "prob") %>%
  bind_cols(pmd_test)

pmd_prediction4 %>%
  roc_curve(truth = violent, .pred_Yes) %>%
  autoplot()
```
]
.pull-right[
- The area under the curve is 0.496 showing again there is no correlation between the sex of the prisoner and the violence of the crimes they commit 
]

---
# Summary of Findings

- The models show us that the chance of reoffence cannot be predicted by any variable, including the violence of the crimes committed.
---
class: inverse, middle, centre

# Are there any questions?

---
